---
title: "hw2"
author: "lin du ld2770"
date: "September 17, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data("ToothGrowth")
VC = ToothGrowth[ToothGrowth[,'supp'] == 'VC','len']
OJ = ToothGrowth[ToothGrowth[,'supp'] == 'OJ','len']

hist(OJ)
```
```{r}
qqnorm(OJ)

```
```{r}
hist(VC)
```
```{r}
qqnorm(VC)
```

We can see from graphs anc conclude that OJ is left skewed without outliers, but VC is normal distributed without outliers.

1 a) parametric procedure
```{r}
t.test(VC,OJ)
```

P-value is larger than 0.05, and we can conclude that we can not reject the null hypothesis, which means the two differences are same.

b) Non parametric procedure
```{r}
wilcox.test(VC,OJ)
```

P-value is larger than 0.05, and we can conclude that we can not reject the null hypothesis, which means the two differences are same.

2 Ignore 'dose' and compute the following, based on a suitable bootstrap method:

a)A 95% confidence interval for the difference in median weight for the two groups

```{r}
library('boot')
fun_median <- function(data1,data2,indices) {
return(median(data1[indices]) - median(data2[indices]))
}
diff_median <- boot(OJ, data2 = VC, R = 500, statistic = fun_median)
boot.ci(diff_median, conf = 0.95, type = "bca")
```

b)A 95% bootstrap confidence interval for the ratio of the variances of the two groups
```{r}
func_ratio <- function(data1, data2, indices) {
return(var(data1[indices])/var(data2[indices]))
}
ratio <- boot(OJ, data2 = VC, R = 500, statistic = func_ratio)
boot.ci(ratio, conf = 0.95, type = "bca")
```



3 a)Ignore 'dose', and determine whether there is a significant difference in the proportions of
the two groups classified as "HIGH" using a suitable test and a 95% confidence interval.
```{r}
for (i in 1:nrow(ToothGrowth)){
 if (ToothGrowth[i,'len'] >20){
   ToothGrowth[i,'level'] = 'high'
   
 }
  else {ToothGrowth[i,'level'] = 'low'}
}

VC_level = ToothGrowth[ToothGrowth[,'supp'] == 'VC','level']
OJ_level = ToothGrowth[ToothGrowth[,'supp'] == 'OJ','level']


num_vc = 0
num_oj = 0
for (i in 1:length(VC_level)){
  if (VC_level[i] == "high"){
    num_vc = num_vc + 1
  }
}

for (i in 1:length(OJ_level)){
  if (OJ_level[i] == "high"){
    num_oj = num_oj + 1
  }
}

res <- prop.test(x = c(num_vc,num_oj), n = c(length(VC_level),length(OJ_level)))
res
```

p-value is not small enough, we can't reject the null hypothesis, means the difference in the proportions of the two groups  classified as "HIGH"  is 0.

b)Repeat the above, taking into account "dose", and comment on the results.
```{r}
num_low_0.5=0
num_low_1=0
num_low_2=0
num_high_0.5=0
num_high_1=0
num_high_2=0
for (i in 1:nrow(ToothGrowth)){
  if (ToothGrowth$level[i] == 'low'){
    if (ToothGrowth$dose[i] == 0.5){
      num_low_0.5= num_low_0.5 + 1
    }
     if (ToothGrowth$dose[i] == 1.0) {
       num_low_1 = num_low_1 +1
     }
    if (ToothGrowth$dose[i] == 2.0) {
      num_low_2 = num_low_2 + 1
    }
  }
  else{
    if (ToothGrowth$dose[i] == 0.5){
      num_high_0.5 = num_high_0.5 + 1
    }
    if (ToothGrowth$dose[i] == 1.0){
      num_high_1 = num_high_1 + 1
    }
    if (ToothGrowth$dose[i] == 2.0){
      num_high_2 = num_high_2 + 1
    }
  }
}

dose_level <- matrix(c(num_high_0.5,num_low_0.5,num_high_1,num_low_1,num_high_2,num_low_2), 3,2,byrow=TRUE)
colnames(dose_level) = c('High','Low')
rownames(dose_level) = c('dose_0.5','dose_1.0','dose_2.0')
fisher.test(dose_level)

```

p-value is small enough, we can reject the null hypothesis, means taking into account "dose", the difference in the proportions of the two groups classified as "HIGH" is not 0.
