---
title: "hw6"
author: "lin du ld2770"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
data <- ChickWeight %>%
group_by(Chick) %>%
mutate(birthWeight = weight[Time==0])
data
```


```{r}
ChickWeight_10 <- data[data["Time"]==10,]
ChickWeight_18 <- data[data["Time"]==18,]
ChickWeight_20 <- data[data["Time"]==20,]
```

```{r}

ancova1 <- lm(weight ~ birthWeight + Diet, data=ChickWeight_10)
anova(ancova1)
```


```{r}
ancova2 <- lm(weight ~ birthWeight+ Diet, data=ChickWeight_18)
anova(ancova2)
```



```{r}
ancova3 <- lm(weight ~ birthWeight + Diet, data=ChickWeight_20)
anova(ancova3)
```
For day 10, the p-value is 0.00212 with adjusting for baseline, so there is a significant difference in the mean weights of the four groups on Day 10 at 0.05 level of significance.
For day 18, the p-value is 0.0234 with adjusting for baseline, so there is a significant difference in the mean weights of the four groups on Day 18 at 0.05 level of significance.
For day 10, the p-value is 0.0109 with adjusting for baseline, so there is a significant difference in the mean weights of the four groups on Day 20 at 0.05 level of significance.

ANOVA
```{r}
## using compound symmetry for the variance-covariance structure
library(nlme)
data1 <- data[data$Time %in% c(10,18,20), ]
data1 <- groupedData(weight ~ Diet | Chick, data = data1)
compsym <- gls(weight ~ birthWeight + Diet*factor(Time), data = data1,
corr = corCompSymm(form = ~ 1 | Chick))
anova(compsym)
```
The P-value of Diet is 0.0013 when using compound symmetry for the variance-covariance structure, so
there is a significant difference in the mean weights of four groups on Days 10, 18 and 20, at 0.05 level of
significance.

```{r}
## using unstructured covariance for the variance-covariance structure
## Option "corr = corSymm" specifies that the correlation structure is unstructured
unstruct <- gls(weight ~ birthWeight + Diet*factor(Time), data = data1,
corr=corSymm(form= ~ 1 | Chick),
weights = varIdent(form = ~ 1 | Time))
anova(unstruct)
```
The P-value of Diet is less than 0.0001 when using unstructured covariance for the variance-covariance
structure, so there is a significant difference in the mean weights of four groups on Days 10, 18 and 20, at
0.05 level of significance.
```{r}
anova(compsym,unstruct)
```
The unstructured covariance structure among repeated measures have relatively smaller AIC and BIC statistics.
Te degrees of freedom of unstructured covariance is 19, which is greater than 15 in compound symmetry
structure (the model has increased the number of parameters that had to be estimated), but it could be
outweighed by the improvement of the AIC, BIC and log likelihood statistics. So the repeated measures
ANOVA when assuming unstructured covariance can better fit the data.


3) Tests for the assumptions

Assumptions for ANCOVA
Assumption 1: equality of slopes-interaction is not significiant,
testing the equality of slopes that the covariate is associated with the
outcome the same way between groups we are just interested in the
interaction term here.
```{r}
res1<-aov(weight~ birthWeight + Diet + birthWeight:Diet,data=ChickWeight_10)
summary(res1)
```


Assumption 2: linearity of slopes.
```{r}
par(mfrow=c(1,3))
plot(ChickWeight_10$birthWeight,ChickWeight_10$weight,col=ChickWeight_10$Diet,xlab="Time",ylab="weight")
plot(ChickWeight_18$birthWeight,ChickWeight_18$weight,col=ChickWeight_18$Diet,xlab="Time",ylab="weight")
plot(ChickWeight_20$birthWeight,ChickWeight_20$weight,col=ChickWeight_20$Diet,xlab="Time",ylab="weight")
```


Assumption 3: Equality of the two groups on the covariate.
```{r}
aov(weight~Diet,data=ChickWeight_10)
```
#We can conclude that these three dates are not equal.

Assumption 4: Homogeneity of variance
```{r}
library(car)

leveneTest(weight~Diet,center=mean,data=ChickWeight_10)
```

Assumption 5: check validity for normality

```{r}
shapiro.test(resid(ancova1))
```
```{r}
shapiro.test(resid(ancova2))
```
```{r}
shapiro.test(resid(ancova3))
```

```{r}
par(mfrow=c(2,3))
hist(resid(ancova1)); hist(resid(ancova2)); hist(resid(ancova3))
qqnorm(resid(ancova1)); qqnorm(resid(ancova2)); qqnorm(resid(ancova3))
```
From ANCOVA, the p-value for the Shapiro-Wilk normality test are 0.02894, 0.8441, 0.7082. The first one for
Day10 is less than 0.05, which means that the distribution of the data are significantly different from normal
distribution. The p-values of the second and third (Day 18 and 20) are greater than 0.05, so the distribution
of the data are not significantly different from normal distribution, then can assume the normality.

Check the assumption of ANOVA
```{r}
#check validity for normality
shapiro.test(resid(fit.compsym))
```
```{r}
shapiro.test(resid(fit.unstruct))
```
```{r}
par(mfrow=c(2,2))
hist(resid(fit.compsym)); hist(resid(fit.unstruct))
qqnorm(resid(fit.compsym)); qqnorm(resid(fit.unstruct))
```
For for repeated measures ANOVA, the p-values of the Shapiro-Wilk normality test are 0.02499 and 0.02359,
which means that the distribution of the data are significantly different from normal distributionthe, so
normality assumption does not hold. Also, we can see from the histograms and qq-plots that there is abvious
deviation from normality.
```{r}
library(car)
leveneTest(weight ~ Diet, center = mean, data = ChickWeight_10)
```
```{r}
leveneTest(weight ~ Diet, center = mean, data = ChickWeight_18)
```
```{r}
leveneTest(weight ~ Diet, center = mean, data = ChickWeight_20)
```
```{r}
par(mfrow = c(1, 3))
plot(fitted(ancova1), resid(ancova1), main = "Residual plot", xlab = "Fitted values", ylab = "Residuals")
abline(h = 0, col = "blue")
plot(fitted(ancova2), resid(ancova2), main = "Residual plot", xlab = "Fitted values", ylab = "Residuals")
abline(h = 0, col = "blue")
plot(fitted(ancova3), resid(ancova3), main = "Residual plot", xlab = "Fitted values", ylab = "Residuals")
abline(h = 0, col = "blue")
```
when testing the equal variance assumptions, p-valuesof the Levene's test for homogeneity of variance are
0.2657, 0.3664 and 0.3358, all of them are greater than 0.05, so we can conclude that the assumption of equal
variances is satisfied. From on the residual plots above, the errors seem to have constant variance as the
residuals are scattered randomly around zero.

